---
title: "Week 03 Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Exercises

1. We'll be exploring [fuel economy data](https://www.fueleconomy.gov/feg/download.shtml) from 1984-2021. You can review a data dictionary for the dataset [here](https://www.fueleconomy.gov/feg/ws/index.shtml#vehicle).

Utilize the following chunk to get started with your analysis.

```{r}
library(tidyverse)
library(here)

vehicles <- read_csv(here::here("Data", "vehicles.csv"),
                     col_types = cols(.default = "c"))
```

Describe what is happening in the `col_types` parameter. Why do you think this is necessary?

```{r}
vehicles
```


Cols looks at each of the columns in the input data, the ".default = 'c'" portion is saying that all of the columns should be read as character types. This may be necessary if you are trying to maintain the integrity of all of the values in the dataset and are trying to prevent the data from being manipulated (i.e running the numeric values through calculations.) Or maybe it's for data uniformity.




2. Create a reduced version of the dataset that includes the following fields:

-   `make`
-   `model`
-   `year`
-   `cylinders`
-   `drive`
-   `trany`
-   `evMotor`
-   `fuelCost08`
-   `highway08`
-   `city08`
-   `feScore`

Ensure that the `year`, fuelCost08`, `highway08`, `city08`, and `feScore` fields are coerced to a double.

Assign this to a new dataframe called `vehicles_sub`.

```{r}
vehicles_sub_temp <-  #temporary dataframe before mutation
  select(vehicles,
         make, model, year, cylinders, drive, trany, evMotor, fuelCost08, highway08, city08, feScore)
```

```{r}
vehicles_sub <-  #changing the specified fields to doubles
  mutate(vehicles_sub_temp,
         year = as.double(year),
         fuelCost08 = as.double(fuelCost08),
         highway08 = as.double(highway08),
         city08 = as.double(city08),
         feScore = as.double(feScore))
```



3. Conduct an exploratory analysis of each field in the dataset. This should include:

- Demonstrating the structure of the dataset
- Counts of each character field (in addition to the `year` field)
- Distributions of each numeric field

Indicate anything interesting that you notice.

*Hint: When finding counts of fields, sometimes it is helpful to sort the field from highest to lowest values.*

```{r}
glimpse(vehicles_sub)  
```


```{r}
vehicles_sub %>%
  group_by(make) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))

vehicles_sub %>%
  group_by(model) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))

vehicles_sub %>%
  group_by(year) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))

vehicles_sub %>%
  group_by(cylinders) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))

vehicles_sub %>%
  group_by(drive) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))


vehicles_sub %>%
  group_by(trany) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))

vehicles_sub %>%
  group_by(evMotor) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))
```

```{r}
vehicles_sub %>%
  group_by(fuelCost08) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))


vehicles_sub %>%
  group_by(highway08) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))

vehicles_sub %>%
  group_by(city08) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))

vehicles_sub %>%
  group_by(feScore) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))
```

4. Some of the fields need to be cleaned to ease analysis. 

- Create a new field in `vehicles_sub` called `evMotor_flag` that indicates whether or not the vehicle has an electric motor.

```{r}
mutate(vehicles_sub,
       evMotor_flag <-
         filter(evMotors == "N/A" ))
```


5. Create a graph that demonstrates how median city mileage has changed over time.

```{r}
ggplot(vehicles_sub,
       aes(x = year,
           y = city08)) +
  geom_point()
```

6. Consider the top 5 car makes whose fuel economies have been evaluated most frequently. Create a graph that demonstrates how the city fuel mileage has changed over time for these makes.

```{r}
vehicles_sub %>%  #find the top 5
  group_by(make) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  arrange(desc(count))

ggplot(vehicles_sub) +
  geom_point(aes(x = year,
                 y = city08,
                 color = Chevrolet, Ford, Dodge, GMC, Toyota)) 
```

7. Create a question about the dataset that must be solved by using the `dplyr` verbs - `select`, `mutate`, `summarize`, `arrange`, etc.

-   Indicate what your question is. Consider a question that could be answered best by utilizing facets, multiple geometries, multiple aesthetics, or some combination of those.
-   Create a graph that assists in answering the question.
-   Write a short interpretation of the graph in order to answer your question.

```{r}
#I would like to see if there is a pattern between overall fuel efficiency and the year the car was made, controlling for a specific drive (in this case a 4) where fuel efficiency is represented by a ratio of highway fuel divided by city


mutate(vehicles_sub,
       efficiency = highway08/city08)
vehicles_sub %>%
  select(drive, efficiency, year) %>%
  filter(drive == "4") %>%
  arrange(desc(year))
```


8. Submit the following information about the project you would like to solve for your course project:

*   The dataset(s) you will be utilizing to perform your analysis. This may be Slacked or emailed to me separately from this document. You may include a direct link to the dataset, a file, the API you intend to extract the data from, or the 
*   The business question that you intend to answer through your analysis.
*   A high level description of how you intend to answer the proposed business question.